### 2018-06-29 paper/4-build.tex
Work on S4.1

The build systems overview (\S\ref{sec-background-summary}) highlighted three
distinct approaches to respecting the dependency order. This subsection explores
their properties and possible implementations.

@@

The topological approach pre-computes a linear order, which when followed, ensures
the \hs{build} is correct regardless of the initial \hs{store}. Given a function
from a key to its dependencies, and the output \hs{key}, you can compute the
linear order by first finding the reachable dependencies of \hs{key},
and then computing a topological sort. However, as we have seen
in~\S\ref{sec-deps}, we can only extract dependencies from an applicative task,
which requires the build system to choose \hs{c}~\hs{=}~\hs{Applicative}, ruling
out dynamic dependencies.

\vspace{-2mm}
\subsubsection{Reordering}\label{sec-restarting}

The topological approach has two downsides: it is limited to \hs{Applicative}
build systems and requires a fresh topological sort each time.  So, while the
actions themselves may be incremental (i.e. unnecessary tasks will not be performed),
the pre-processing is not. We can try to incrementalise the topological sort by
storing the topological order between build runs and assume it to
be correct, but if the build discovers it is wrong, fix it up.

This approach requires a way to abort tasks that have failed due to out-of-date
dependencies. It is also not minimal in the sense that a task may start, do some
meaningful work, then abort. However, in the case of an \hs{Applicative} system,
that work is zero.

\vspace{-2mm}
\subsubsection{Recursive}\label{sec-suspending}

An alternative approach, utilised by the \hs{busy} build system
(\S\ref{sec-general-build}), is to simply build dependencies when they are
requested. By combining that with a transient set of which keys have already
been built, you can obtain a minimal build system.

This approach requires that a task may be started, then during that execution
another task will have to be run. Assuming an IO-driven task structure,
that requires suspending a running task, which can be done with cheap
green threads and blocking (the original approach of \Shake) or using
continuation-passing style (what \Shake does currently). An alternative approach to
suspending a task is to abort it and restart it again later, at the cost
of doing additional work.

>>>>

The build systems overview (\S\ref{sec-background-summary}) highlighted three
distinct task schedulers respecting the dependency order. This subsection
explores their properties and possible implementations.

@@

The topological scheduler pre-computes a linear order, which when followed,
ensures the build result is correct regardless of the initial store. Given a
task description and the output \hs{key}, you can compute the linear order by
first finding the (acyclic) graph of the \hs{key}'s reachable dependencies, and
then computing a topological sort. However, as we have seen in~\S\ref{sec-deps},
we can only extract dependencies from an applicative task, which requires the
build system to choose \hs{c}~\hs{=}~\hs{Applicative}, ruling out dynamic
dependencies.

\vspace{-2mm}
\subsubsection{Restarting}\label{sec-restarting}

The topological scheduler has two downsides: it is limited to \hs{Applicative}
build systems and requires a fresh topological sort each time. So, while the
actions themselves may be incremental (i.e. unnecessary tasks will not be
performed), the pre-processing is not. We can try to incrementalise the
topological sort by storing the topological order between build runs and assume
it to be correct, fixing it up if the build discovers any inconsistency, as
\Excel does.

This approach requires a way to abort tasks that have failed due to out-of-date
dependencies. It is also not minimal in the sense that a task may start, do some
meaningful work, and then abort. However, in the case of an \hs{Applicative}
system, that work is zero.

\Bazel's restarting scheduler does not store the topological order between build
runs; instead, it stores the most recent task dependency information. Since this
information may become outdated, \Bazel may also need to abort a task if a
newly discovered dependency is out-of-date.

\vspace{-2mm}
\subsubsection{Suspending}\label{sec-suspending}

An alternative approach, utilised by the \hs{busy} build system
(\S\ref{sec-general-build}) and \Shake, is to simply build dependencies when
they are requested, suspending the currently running task. By combining that
with tracking the keys that have already been built, one can obtain a minimal
build system with dynamic dependencies.

This approach requires that a task may be started and then suspended until
another task is complete. This can be done with cheap green threads and blocking
(the original approach of \Shake) or using continuation-passing style (what
\Shake does currently).

% An alternative approach to suspending a task is to abort
% it and restart it again later, at the cost of doing additional work.
